Practical Machine Learning Class
HAR Clustering Model/Analysis using Random Forest
=========================================================

To start with. the training data was partititioned into two - training (70%) and cross validation (30%). The partitioned training set was used to create the model and was validated using the cross validation set. Once the model parameters were satisfactory, the model was derived based on the full training set and this derived model was used to predit the test set.

Load the traning/test data

```{r}
setwd("~/My Training/PML")

currentTrainPath = 'train.RData'
currentTestPath = 'test.RData'

training = F

if (training)
{
  trn = read.csv('pml_trn.csv',header=T)
} else {
  modFit<-readRDS(currentTrainPath)
  tst = read.csv('pml_tst.csv',header=T)
}

```

Data Cleanup and Feature Addition
==================================

```{r}
if (training) {
  data<-trn
} else {
  data<-tst
}
  
data<-data[which(data$new_window=="no"),]
head(data)


```

Delete all columns containing NA
================================
```{r}

rmcol<-array()
j<-1
for (i in 1:ncol(data)) {
  if ((sum(is.na(data[,i])) > 0) | (sum(data[,i] == "") > 0)) {
    rmcol[j]<-i
    j<-j+1
  }
}

data<-data[,-rmcol]
rm(rmcol)

tm<-unlist(lapply(strsplit(as.character(data$cvtd_timestamp)," "), function(x) x[2]))
time<-(lapply(strsplit(tm,":"),function(x) {return (round((as.numeric(x[1])) + round(as.numeric(x[2])/60)))}))

data$cvtd_timestamp = factor(NA,levels=c('midnight','morning','noon','afternoon','evening'))

data$cvtd_timestamp[time>22 | time<=6] <- 'midnight'
data$cvtd_timestamp[time>6 & time<=10] <- 'morning'
data$cvtd_timestamp[time>10 & time <= 13] <- 'noon'
data$cvtd_timestamp[time>13 & time <= 18] <- 'afternoon'
data$cvtd_timestamp[time>18 & time <= 22] <- 'evening'

rm(tm,time)

```

Train the Model
===============
```{r}
library(caret)


nam<-names(data)
valCols<-c(which(nam=="accel_dumbbell_x"),which(nam=="magnet_dumbbell_x"),which(nam=="magnet_dumbbell_y"),which(nam=="magnet_dumbbell_z"),which(nam=="roll_forearm"),which(nam=="pitch_forearm"),which(nam=="yaw_forearm"),which(nam=="total_accel_forearm"),which(nam=="gyros_forearm_x"),which(nam=="gyros_forearm_y"),which(nam=="gyros_forearm_z"),which(nam=="accel_forearm_x"),which(nam=="accel_forearm_y"),which(nam=="accel_forearm_z"),which(nam=="magnet_forearm_x"),which(nam=="magnet_forearm_y"),which(nam=="magnet_forearm_z"),which(nam=="classe"))

if (training) {
   data_b<-data
   values<-data_b$classe
   modFit<-train(classe~.,data<-data_b[,valCols],method="rf")
   prediction<-data$classe
   saveRDS(modFit,currentTrainPath)
   error<-sum(values!=prediction)/length(values)
   paste("Training Set Error:",(error)," ")
   
} else {
  values<-data$classe
  prediction<-predict(modFit,data[,valCols])
  error<-sum(values!=prediction)/length(values)
  paste("Test Set Error:",(error)," ")
}
```
